---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hoaddata

<!-- badges: start -->
[![update-data.yaml](https://github.com/njahn82/hoaddashtest/actions/workflows/update-data.yaml/badge.svg)](https://github.com/njahn82/hoaddashtest/actions/workflows/update-data.yaml)

<!-- badges: end -->

This package contains indicators about the open access uptake of
hybrid journals, which belong to national transformative agreements in Germany
as listed by the 
[Open Access Monitor](https://open-access-monitor.de/#/publications).

## Installation

You can install hoaddata from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("njahn82/hoaddashtest", dependencies = "Imports")
```
## Data tables

This package provides the following data tables.

### `oam_hybrid_jns`

Hybrid Journals listed in the Open Access Monitor. Data were 
gathered from <https://doi.org/10.26165/JUELICH-DATA/VTQXLM> and enriched with
ISSN variants.

```{r}
oam_hybrid_jns
```

### `cc_jn_ind`
Aggregated data about the prevalence of Creative Commons 
license variants by year and hybrid journal as obtained from Crossref.

```{r}
cc_jn_ind
```

### `cr_olax_inst`

Article-level affiliation data from first authors as obtained
from OpenAlex.

```{r}
cr_olax_inst
```


## Data re-use and licenses

Datasets are released into the public domain.

Anyone is free to copy, modify, publish, use, compile, sell, or distribute these materials in any form, for any purpose, commercial or non-commercial, and by any means.

Crossref asserts no claims of ownership to individual items of bibliographic metadata and associated Digital Object Identifiers (DOIs) acquired through the use of the Crossref Free Services. Individual items of bibliographic metadata and associated DOIs may be cached and incorporated into the user's content and systems.

This work re-used the following dataset: 

Pollack, Philipp; Lindstrot, Barbara; Barbers, Irene, 2021, "Open Access Monitor: Zeitschriftenlisten", <https://doi.org/10.26165/JUELICH-DATA/VTQXLM>.

published under CC BY 4.0.
