---
title: "Overview: Hybrid OA data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview: Hybrid OA data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hoaddata)
```

## Datasets

This package provides the following data tables.

### `oam_hybrid_jns`

Hybrid Journals listed in the Open Access Monitor. Data were 
gathered from <https://doi.org/10.26165/JUELICH-DATA/VTQXLM> and enriched with
ISSN variants.

```{r}
oam_hybrid_jns
```

### `cc_jn_ind`
Aggregated data about the prevalence of Creative Commons 
license variants by year and hybrid journal as obtained from Crossref.

```{r}
cc_jn_ind
```

### `cc_openalex_inst_jn_ind`

First author country affiliations per journal, 
year and Creative Commons license

```{r}
cc_openalex_inst_jn_ind
```

### `cc_openalex_inst`

Article-level affiliation data from first authors as obtained
from OpenAlex.

```{r}
cc_openalex_inst
```
