---
title: "Overview: Hybrid OA data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Overview: Hybrid OA data}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hoaddata)
library(dplyr)
```

## Datasets

This package provides the following data tables.

### `oam_hybrid_jns`

Hybrid Journals listed in the Open Access Monitor. Data were gathered
from <https://doi.org/10.26165/JUELICH-DATA/VTQXLM> and enriched with
ISSN variants.

```{r}
oam_hybrid_jns
```

Number of investigated journals:
`r length(unique(oam_hybrid_jns$cr_journal_id))`

By agreement

```{r}
oam_hybrid_jns %>%
  group_by(agreement, lead) %>%
  summarise(journals = n_distinct(cr_journal_id)) %>%
  arrange(desc(journals))
```

### `oam_jn_ind`

Aggregated data about the prevalence of Creative Commons license
variants by year and hybrid journal as obtained from Crossref.

```{r}
oam_jn_ind
```

Number of active journals: `r length(unique(oam_jn_ind$cr_journal_id))`

Number of open access articles with Creative Commons license:
`r oam_jn_ind %>% filter(!is.na(cc)) %>% .$cc_total %>% sum()`

Creative Commons Breakdown:

```{r}
oam_jn_ind %>%
  filter(!is.na(cc)) %>%
  group_by(cc) %>%
  summarise(articles = sum(cc_total)) %>%
  arrange(desc(articles))
```

### `oam_jn_aff`

First author country affiliations per journal, year and Creative Commons
license

```{r}
oam_jn_aff
```

### `oam_cc_inst`

Article-level affiliation data from first authors as obtained from
OpenAlex.

```{r}
oam_cc_inst
```

```{r}
oam_cc_inst %>%
  group_by(cc) %>%
  summarise(articles = n_distinct(doi))
```

### `oam_cr_md`
 
Crossref metadata coverage.

```{r}
oam_cr_md
```

